Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations

### 2020-04-29 paper/tidy-missing-data-paper.Rmd
finish of Overview

All missings can be displayed using a heatmap style visualization. This is achieved using `vis_miss()` from the \pkg{visdat} package [@visdat], which also provides summaries of missingness overall in the legend, and for each column (Figure \@ref(fig:vis-miss)). The user can also apply clustering to the rows, and arrange columns by missingness (Figure \@ref(fig:vis-miss)B). Similar visualizations are available in other missing data packages such as \pkg{VIM}, \pkg{mi}, \pkg{Amelia}, and \pkg{MissingDataGUI}. A key improvement is `vis_miss()` orients the visualization analogous to a regular data structure: variables form columns in the visualization and are named at the top, and each row is an observation. Using \pkg{ggplot2} makes the plot easily customizable.

@@

The number of times certain variables go missing together can be visualized using an "upset plot" [@Conway2017]. Similar to a Venn diagram, an upset plot shows the size and features of overlapping sets, but scales well with more variables. Figure \@ref(fig:airquality-upset) shows only Ozone and Solar.R have missing values. The bottom right show the combinations of missingness, the bottom left showing missingness in each variable, and the top showing the combinations of missingness. The bottom left shows Ozone has the most missing values, and the top shows 2 cases where both Solar.R and Ozone go missing together.

>>>>

Both variables and cases can be displayed using a heatmap style visualization,available with `vis_miss()` from the \pkg{visdat} package [@visdat]. This also provides numerical summaries of missingness in the legend, and for each column (Figure \@ref(fig:vis-miss)). The user can also apply clustering to the rows, and arrange columns by missingness (Figure \@ref(fig:vis-miss)B). Similar visualizations are available in other missing data packages such as \pkg{VIM}, \pkg{mi}, \pkg{Amelia}, and \pkg{MissingDataGUI}. A key improvement is `vis_miss()` orients the visualization analogous to a regular data structure: variables form columns in the visualization and are named at the top, and each row is an observation. Using \pkg{ggplot2} makes the plot easily customizable. This also uses the shadow matrix as the basis for the plot.

@@

The number of times certain variables go missing together can be visualized using an "upset plot" [@Conway2017]. An alternative to a Venn diagram, an upset plot shows the size and features of overlapping sets, but scales well with more variables. Figure \@ref(fig:airquality-upset) shows the airquality data, and illustrates  that only Ozone and Solar.R have missing values. The bottom right show the combinations of missingness, the bottom left showing missingness in each variable, and the top showing the combinations of missingness. The bottom left shows Ozone has the most missing values, and the top shows 2 cases where both Solar.R and Ozone go missing together. The upset plot is constructed from the shadow matrix.
